The ChEMBL database as Linked Open Data

### 2013-01-14 article.tex
Extended the descriptions of types and the drug role more verbosely

The ChEMBL database
recognizes various types: \verb+pro:PR_0000001+, \verb+chembl:ADMET+, \verb+chembl:CELL-LINE+,
\verb+chembl:NUCLEIC-ACID+, \verb+chembl:ORGANISM+, \verb+chembl:SUBCELLULAR+, \verb+chembl:TISSUE+,
\verb+chembl:UNCHECKED+, and \verb+chembl:UNKNOWN+. The latter two are currently defined as
explicit types, thus effectively implementing a closed-world approach.

@@

For drug discovery, the drug-like compounds themselves are the main topic of study.
ChEMBL contains many different compound types, mostly small molecules,
but also peptides, proteins, antibodies, oligosaccharides, oligonucleotides, and
even cells. ChEMBL-RDF follows the approach used by the CHEMINF ontology,
and encodes drugs as classes, rather than instances, and they subclass
other classes defined by several ontologies for these types. For example, proteins subclass the
Protein class from the PRotein Ontology (PR\_000000001), small
molecules subclass the Chemical Entity class from the CHEMINF
ontology (CHEMINF\_000000) (which is itself equivalent to the Chemical Entity class in ChEBI, CHEBI\_24431), 
and oligosaccharides and oligonucleotides
subclass their respective matches in the CHEBI ontology (CHEBI\_50699
and CHEBI\_7754). To ensure we can look up all entities for which activities
are reported, we subclass everything from the BFO 1.1 class MaterialEntity~\cite{Smith2004},
which is a superclass of the CHEMINF Chemical Entity class.

It should also be noted that the relationship of these structural classes 
is that of having a \textit{role} of drug, rather than a subclass relationship.
Moreover, the triples only attach this drug role to approved drugs, as defined by the
ChEMBL database. 
This is triplified using the OBO and ChEBI ontologies in the following manner, where CHEBI\_23888 is the drug role:

>>>>

The ChEMBL database
recognizes various target types: \verb+pro:PR_0000001+, \verb+chembl:ADMET+, \verb+chembl:CELL-LINE+,
\verb+chembl:NUCLEIC-ACID+, \verb+chembl:ORGANISM+, \verb+chembl:SUBCELLULAR+, \verb+chembl:TISSUE+,
\verb+chembl:UNCHECKED+, and \verb+chembl:UNKNOWN+. The latter two are currently defined as
explicit types, thus effectively implementing a closed-world approach.

@@

For drug discovery, the drug-like compounds themselves are the main topic of study.
ChEMBL contains many different compound types, mostly small molecules,
but also peptides, proteins, antibodies, oligosaccharides, oligonucleotides, and
even cells. ChEMBL-RDF follows the approach used by the CHEMINF ontology,
and encodes drugs as classes, rather than instances, and they subclass
other classes defined by several ontologies for these types. For example, protein drugs subclass the
Protein class from the PRotein Ontology (PR\_000000001), instead that these drugs are
of the Protein type. Each drug is a class itself. Likewise, small
molecules subclass the Chemical Entity class from the CHEMINF
ontology (CHEMINF\_000000) (which is itself equivalent to the Chemical Entity class in ChEBI, CHEBI\_24431), 
and oligosaccharides and oligonucleotides
subclass their respective matches in the CHEBI ontology (CHEBI\_50699
and CHEBI\_7754). To ensure we can look up all entities for which activities
are reported, we subclass everything from the BFO 1.1 class MaterialEntity~\cite{Smith2004},
which is a superclass of the CHEMINF Chemical Entity class.

It should also be noted that all drug-like entities in ChEMBL are not ontologically
subclassing a drug class. Instead, we give the entities the role of drug: only within a
particular context, they would be drug. If not used as medication, they are not drug.
Moreover, the triples only attach this drug role to approved drugs, as defined by the
ChEMBL database. All drug-like entities in the database do not have the role of drug.
The role is triplified using the OBO and ChEBI ontologies in the following manner,
where CHEBI\_23888 is the ontological entry for of ``drug role'' in ChEBI:
